---
title: "Experimentalpraktikum"
subtitle: "EEG Analyse in R"
author: "Joshua Ziegler"
date: "10.02.2021"
output: 
  revealjs::revealjs_presentation:
    self_contained: false
    reveal_plugins: ["menu"]
    transition: slide 
---
<style>
    .reveal .slides {
        text-align: left;
    }
    .reveal .slides section>* {
        margin-left: 0;
        margin-right: 0;
    }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(eegUtils) 
library(tidyverse)
```

# Was ist R? 
# Warum R?
# eegUtils
# Studie
# Auswertung
## Laden der Rohdaten
<small>eegUtils-Funktion: ```import_raw```</small>
```{r, echo=T, results='hide', message=F, eval=FALSE}
# Liste der vhdr-Dateien im "raw"-Ordner
temp <- list.files(path = "./raw", pattern="*.vhdr")
# Leere Liste erstellen
DORlist <- list()
# Daten importieren
for (i in temp){
  filename <- paste0(i)     # filename from temp
  DORlist[[i]] <-           # save the following elements in a list
    assign(filename,        # assign respective filename to data
           import_raw(i,    # import data set
                      file_path = "./raw",   # path to subdirectory "raw"
                      participant_id = make.names(gsub("*.vhdr$", "", i)))  # set particpant ID
  )
  rm(list = ls(pattern = "vhdr$"))  # remove files that are not in list
}
```

## Überblick über die Daten
browse_data (screenshot) & Listenorganisation (screenshot)

## Preprocessing
<small>eegUtils-Funktionen: ```electrode_locations, eeg_filter, eeg_reference, epoch_data```</small>
```{r, echo=T, results='hide', message=F, eval=FALSE}
# Preprocessing
DORlist <-      # overwrite list
  lapply(DORlist, function(x){     # apply the following to every element in list
# ELECTRODE LOCATIONS
   x <- electrode_locations(x,
                           overwrite = T,
                           method = "biosemi64")
# FILTER
   x <- eeg_filter(x,
                   low_freq = 0.1,
                   high_freq = 30,
                   method = "fir")
# RE-REF
   x <- eeg_reference(x, 
                      ref_chans = c("A1", "A2"))
# EPOCH
   x <- epoch_data(x, 
                   events = c("S111",
                              "S112",
                              "S212",
                              "S211",
                              "S121",
                              "S123",
                              "S223",
                              "S221"), 
                   epoch_labels = c("oeStan",    # o-e (o-Standard)
                                    "eoStan",    
                                    "oeDev",
                                    "eoDev",     # e-o (o-Deviant)
                                    "ouStan",    # o-u (o-Standard)
                                    "uoStan",    # u-o (u-Standard)
                                    "ouDev",     # o-u (u-Deviant)
                                    "uoDev"),    # u-o (o-Deviant)
                   time_lim = c(-.2, .8),
                   baseline = c(-.2, 0))
}
)
```

## ICA
<small>eegUtils-Funktion: ```run_ICA```</small>
```{r, echo=T, eval=FALSE}
# Leere ICA Liste
ICAlist <- list()
# ICA
ICAlist <- 
  lapply(DORlist, function(x){
           run_ICA(x)
    }
    )
```

## Überblick ICA
<small>eegUtils-Funktion: ```view_ica```</small>
(sceenshot)

## ICA Komponenten
<small>eegUtils-Funktionen: ```ar_eogcor, ar_acf, ar_chanfoc, ar_trialfoc, apply_ica```</small>
```{r, echo=T, eval=FALSE}
# ICA components (using map2() from purrr package (tidyverse))
DORlist <-
  map2(DORlist, ICAlist, function(x, y){
    # print participant id
    print(x$epochs$participant_id[1])
    # ICA EOG
    print("ICA EOG")
    ICA_comp_eog <- ar_eogcor(decomp = y, 
                              data = x, 
                              HEOG = c("HEOGli", "HEOGre"), 
                              VEOG = c("VEOGo", "VEOGu"),
                              plot = F)
    # ICA MUSCLE
    print("ICA Muscle")
    ICA_comp_mus <- ar_acf(y,
                           plot = F)
    # ICA CHANNEL
    print("ICA Channel")
    ICA_comp_chn <- ar_chanfoc(y,
                               plot = F)
    # ICA TRIAL
    print("ICA Trial")
    ICA_comp_trl <- ar_trialfoc(y,
                                plot = F)
    # COMBINE COMPONENTS    
    ICA_comp_rem <- c(ICA_comp_eog, ICA_comp_mus, ICA_comp_chn, ICA_comp_trl)
    print(ICA_comp_rem)
    # REMOVE COMPONENTS
    x <- apply_ica(data = x,    
                   decomp = y, 
                   comps = ICA_comp_rem)
  })
```

```{r, eval=FALSE}
# REMOVE ICAlist
rm(ICAlist)
```

## Datenanalyse
```{r, echo=T, eval=FALSE}
# CONVERT TO DATA FRAME
DOR_df_FCz <- 
  map(DORlist, function(x){
    x %>%
      filter(epoch_labels %in% c("uoDev", "ouStan", "uoStan", "ouDev")) %>%
      as.data.frame(long = T) %>%
      filter(electrode == "FCz")
  })
# COMBINE DATASETS
DOR_gravg_FCz <- bind_rows(DOR_df_FCz)
```

head?

## Plots
## ggplot Code
```{r, echo=T, eval=FALSE}
# Plot o Data
o_diff_plot <-
DOR_gravg_FCz %>%
  filter(epoch_labels %in% c("uoDev", "ouStan")) %>%
  ggplot(aes(x = time, y = amplitude)) +
    stat_summary(fun = mean, 
               geom = "line", 
               aes(colour = epoch_labels)) + 
    facet_wrap(~electrode) + # wenn mehrere elektroden
    scale_y_reverse() + 
    theme_light() +
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0)
```

## /o:/
```{r, eval=FALSE}
print(o_diff_plot)
```

## /u:/
```{r, eval=FALSE}
u_diff_plot <- 
DOR_gravg_FCz %>%
  filter(epoch_labels %in% c("ouDev", "uoStan")) %>%
  ggplot(aes(x = time, y = amplitude)) +
  stat_summary(fun = mean, 
               geom = "line", 
               aes(colour = epoch_labels)) + 
  facet_wrap(~electrode) + # wenn mehrere elektroden
  scale_y_reverse() + 
  theme_light() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0)

print(u_diff_plot)
```


## iMMN
```{r, echo=T, eval=FALSE}
# Mittelwert für jeden Zeitpunkt, aufgeteilt nach Bedingung
diff.avg <-
  DOR_gravg_FCz %>%
  group_by(time, epoch_labels, electrode, participant_id) %>%
  summarise_at("amplitude", mean) %>%
  ungroup()

# Differenzen (Dev-Stan) bilden
# long to wide format
diff.avg <-
  diff.avg %>%
  pivot_wider(names_from = epoch_labels,
              values_from = amplitude)
  
# Zeilen substrahieren
diff.avg <-
  diff.avg %>%
  mutate(o_diff = uoDev - ouStan,
         u_diff = ouDev - uoStan,
         .keep = "unused")

# wide to long format
diff.avg <-
  diff.avg %>%
  pivot_longer(cols = c(o_diff, u_diff), 
               names_to = "epoch_diff", 
               values_to = "amplitude_diff")
```

## Plot
```{r, echo=T, eval=FALSE}
# Plot differences
immn.plot <-
diff.avg %>%
  ggplot(aes(x = time, y = amplitude_diff)) +
  stat_summary(fun = mean, 
               geom = "line", 
               aes(colour = epoch_diff)) + 
  facet_wrap(~electrode) + # wenn mehrere elektroden
  scale_y_reverse() + 
  theme_light() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0)

print(immn.plot)
```

## Inferenzstatistik
```{r, echo=T, eval=FALSE}
# AVERAGE DATA OVER 120-200 ms
diff.avg.crit <-
  diff.avg %>%
    filter(time >= 0.12,  # 120 ms
          time <= 0.2)    # 200 ms

# mean by participant and epoch
diff.avg.crit <-
  diff.avg.crit %>% 
  group_by(electrode, participant_id, epoch_diff) %>%  # keep electrode, participant & epoch
  summarise_at("amplitude_diff", mean) %>%             # mean amplitude
  ungroup()

# REPEATED MEASURES ANOVA ON DIFFERENCES
print(summary(aov(amplitude_diff ~ epoch_diff + Error(participant_id), 
            data = diff.avg.crit)))
```

# Diskussion

# Fazit
- Pros:
- Cons:
